My title

### 2024-11-22 paper/paper.qmd
update paper data section

library(tidyverse)
library(here)
library(arrow)
library(corrplot)
library(knitr)

raw_data_climate <- read_csv(here("data/01-raw_data/climateyvr.csv"))
raw_data_ahccd <- read_csv(here("data/01-raw_data/ahccdyvr.csv"))
analysis_data <- read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))
m1 <- readRDS(here("models/mlr_1.rds"))
m2 <- readRDS(here("models/mlr_2.rds"))
glm_nolog <- readRDS(here("models/glm_nolog.rds"))
glm <- readRDS(here("models/glm_model.rds"))
brm <- readRDS(here("models/brm_model.rds"))
```

@@

With the data obtained from @canadian_centre_for_climate_services_adjusted_2022 and @meteorological_service_of_canada_past_2023, this research paper aims to identify the factors influencing temperature at Vancouver International Airport. Located on the west coast of Richmond, the airport sits on Sea Island, surrounded by water. As a transportation hub for passengers and freight, it is important to assess the location's safety in a warming climate.

@@

# Data {#sec-data}
```{r}

selected_data_1 <- analysis_data[, c("wind_speed", "pressure_station", "snow", "log_mean_temp", "total_rain", "gust_speed_km_h")]
pairs(as.data.frame(selected_data_1), main = "Pairwise Scatterplots (Excluding Max Temp, and Min Temp)")
cor_matrix <- cor(selected_data_1)
print(cor_matrix)
corrplot(cor_matrix, method = "color", type = "upper", 
         tl.col = "black", tl.srt = 45, 
         col = colorRampPalette(c("blue", "white", "red"))(200))

selected_data_3 <- analysis_data[, c("wind_speed", "pressure_station", "log_mean_temp", "total_rain", "gust_speed_km_h")]
pairs(as.data.frame(selected_data_3), main = "Pairwise Scatterplots of log transformed temperature (Excluding Snow, Max Temp, and Min Temp)")
cor_matrix <- cor(selected_data_3)
print(cor_matrix)
corrplot(cor_matrix, method = "color", type = "upper", 
         tl.col = "black", tl.srt = 45, 
         col = colorRampPalette(c("blue", "white", "red"))(200))

@@

@@

## Data Cleaning

###

@@

```{r}
fitted_values <- fitted(m1)
ggplot(data = analysis_data, aes(x = fitted_values, y = mean_temp_F)) +
  geom_point(color = "blue", alpha = 0.7) +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed", size = 1) +
  labs(x = "Predicted Temperature", y = "Actual Temperature", title = "Actual Temperature vs. Predicted Values")

@@

```{r}
fitted_values <- fitted(m2)
ggplot(data = analysis_data, aes(x = fitted_values, y = mean_temp_F)) +
  geom_point(color = "blue", alpha = 0.7) +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed", size = 1) +
  labs(x = "Predicted Temperature", y = "Actual Temperature", title = "Actual Temperature vs. Predicted Values")
```

@@

\appendix

# Appendix {#sec-appendix}

@@

>>>>

library(tidyverse)
library(here)
library(arrow)
library(corrplot)
library(knitr)
library(kableExtra)

raw_data_climate <- read_csv(here("data/01-raw_data/climateyvr.csv"))
raw_data_ahccd <- read_csv(here("data/01-raw_data/ahccdyvr.csv"))
analysis_data <- read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))
train_data <- read_parquet(here("data/02-analysis_data/train_data.parquet"))
test_data <- read_parquet(here("data/02-analysis_data/test_data.parquet"))
m1 <- readRDS(here("models/mlr_1.rds"))
m2 <- readRDS(here("models/mlr_2.rds"))
glm_nolog <- readRDS(here("models/glm_nolog.rds"))
glm <- readRDS(here("models/glm_model.rds"))
brm <- readRDS(here("models/brm_model.rds"))
```

@@

Temperatures significantly impact airport operations. Rising temperatures significantly affect aircraft performance, potentially leading to take-off weight restrictions and the need for longer runways. This directly impacts airport capacity and operations [@coffel_climate_2015]. Temperature forecast models vary in different locations, and different regions have unique climate characteristics that models may not fully capture [@american_meteorological_society_weather_nodate].

This research paper aims to identify the factors influencing temperature at Vancouver International Airport and build a model for temperature prediction with the data obtained from @canadian_centre_for_climate_services_adjusted_2022 and @meteorological_service_of_canada_past_2023. Located on the west coast of Richmond, the airport sits on Sea Island, surrounded by water. As a transportation hub for passengers and freight, it is important to assess the location's safety in a warming climate.

@@

The data gathering and analysis is done in R [@citeR] with the following packages: knitr [@knitr], tidyverse [@tidyverse], ggplot2 [@ggplot2], dplyr [@dplyr], arrow [@arrow], here [@here], and lubridate [@lubridate].

# Data {#sec-data}

<!-- ```{r}
pairs(as.data.frame(selected_data_0), main = "Pairwise Scatterplots (Including Max Temp, and Min Temp)"),
pairs(as.data.frame(selected_data_1), main = "Pairwise Scatterplots (Excluding Max Temp, and Min Temp)")
``` -->

@@

The limitations are documented in @sec-methodology.

@@

## Data Cleaning {#sec-datacleaning}

The data cleaning process consists of two steps. First, we standardize and clean the column headers. Second, we merge the two datasets into a single combined dataset. The dataset used in this analysis combines information from two distinct sources: climate data (raw_data_climate) and historical weather data (raw_data_ahccd). The analysis spans data collected between 1959 and 2010 for training and testing purposes.

The cleaned dataset contains a range of weather variables providing detailed monthly observations. The `date` variable represents the observation month, standardized to the first day of each month. **`wind_speed`** (km/h) captures average monthly wind speeds, while **`total_precipitation`** (mm) measures the total monthly precipitation, including rain and snow. **`snow`** (mm) records total snowfall, and **`pressure_station`** (kPa) indicates atmospheric pressure at the observation station. **`max_temp`** (°C), **`min_temp`** (°C), and **`mean_temp`** (°C) represent the monthly averages of maximum, minimum, and overall temperatures, respectively. **`total_rain`** (mm) focuses solely on rainfall amounts, distinct from snowfall. **`gust_speed_km_h`** (km/h) records the monthly average of maximum gust speeds. Additionally, constructed variables include **`mean_temp_F`**, the mean temperature was converted to Fahrenheit using 
$$
(\text{mean\_temp} \times 1.8) + 32
$$
, and **`log_mean_temp`**, the log-transformed Fahrenheit temperature, was calculated as 
$$
\log(\text{mean\_temp\_F})
$$
.
All column names were cleaned and standardized using `janitor` in @tidyverse to ensure consistency and readability. Dates were parsed into a unified format (`yyyy-mm-dd`) and aligned with monthly observations using the @lubridate package. The datasets were merged into a single combined dataset using the `date_time` variable as the common key. Finally, constructed variables, including `mean_temp_F` and `log_mean_temp`, were added to facilitate advanced modeling techniques.

## Cleaned Data

The top 6 rows of the cleaned data is displayed in @tbl-cleaned-data.

```{r}
#| label: tbl-cleaned-data
#| tbl-cap: Sample of Cleaned Weather Data
#| echo: false
#| warning: false
analysis_data |>
  head(6) |>
  mutate(log_mean_temp = round(log_mean_temp, 2)) |>
  kable(
    col.names = c("Date", "Wind Speed", "Total Precipitation", "Snow", "Pres.", "Max Temp", "Min Temp", "Mean Temp", "Rain", "Max Gust Speed", "Mean Temp in F", "Log of Mean Temp"),
    booktabs = TRUE
    ) 
```

The summary statistics of the combined dataset is displayed in @tbl-summary.

## Characteristics in Cleaned Data
### Skewness in Mean Temperature

```{r}
#| label: fig-hist-temp
#| fig-cap: Mean Temp Shows More Normality and Less Skewness After Adjustment
#| fig-subcap: ['Original Mean Temp has Skewness and Negative Numbers', 'Mean Temp in F Transformed the Value to All Positive', 'Log-Transformed Data Shows a More Symmetric and Less Skewed Distribution']
#| echo: false
#| warning: false
#| layout-ncol: 2
attach(analysis_data)

hist(mean_temp, 
     breaks = 100, 
     main = "Distribution of Mean Temperature (°C)", 
     xlab = "Mean Temperature (°C)", 
     col = "red")

# Histogram for Mean Temperature in Fahrenheit
hist(mean_temp_F, 
     breaks = 100, 
     main = "Distribution of Mean Temperature (°F)", 
     xlab = "Mean Temperature (°F)", 
     col = "green")

# Histogram for Log of Mean Temperature in Fahrenheit
hist(log_mean_temp, 
     breaks = 100, 
     main = "Log-Transformed Mean Temperature (°F)", 
     xlab = "Log of Mean Temperature (°F)", 
     col = "blue")
```

@fig-hist-temp displays the histogram of the response variable Mean Temperature. @fig-hist-temp-1 shows the original mean temperature, which is skewed and includes negative values, making it unsuitable for direct modeling. To address this, We first transformed the data to Fahrenheit in @fig-hist-temp-2, shifting all values to be positive. However, to further normalize the distribution and reduce skewness, we applied a logarithmic transformation in @fig-hist-temp-3. The log transformation stabilizes variance, improves symmetry, and addresses non-linearity in the data, making it more appropriate for modelling.

### Total Snow Is Zero-Inflated

@fig-boxplot-snow clearly shows significant zero inflation, with a large number of observations concentrated at zero and a few extreme outliers far above the majority of the data. This distribution suggests that the variable snow contains excessive structural zeros, likely representing instances where no snowfall occurred.

<!-- Poor Model Fit: Traditional regression models assume a continuous or discrete distribution of values. The extreme concentration of zeros violates this assumption, leading to biased or inaccurate parameter estimates.
Misleading Interpretability: The effect of snow on the response variable might be confounded by the disproportionate impact of the zeros, making the interpretation of results unreliable.
Violated Assumptions: Many models assume homoscedasticity (constant variance) and normality of residuals. The zero inflation in snow increases heteroscedasticity and skews the residuals.
My bayesian model depends on normal likelihood, so normality and constant variance matters. Gaussian likelihood in your model assumes continuous, symmetric data -->

```{r}
#| label: fig-boxplot-snow
#| fig-cap: Total Snow shows Zero Inflation
#| echo: false
#| warning: false
boxplot(analysis_data$snow,
        main = "Boxplot of Total Snow",
        ylab = "Snow (in cm)", 
        col = "lightblue",
        notch = TRUE)
```

### Maximum Temperature, Minimum Temperature and Mean Temperature Has Strong Linear Relationships

```{r}
#| label: fig-corr-temp
#| fig-cap: Temperature Values Have High Correlations
#| fig-subcap: ['High Correlations Between Max, Min and Mean Temperature', 'Other Predictors Does Not Have High Correlations', 'Linear Relationship Between Max and Mean Temperature', 'Linear Relationship Between Min and Mean Temperature']
#| echo: false
#| warning: false
#| layout-ncol: 2
selected_data_0 <- analysis_data[, c("wind_speed", "pressure_station", "snow",  "min_temp", "log_mean_temp", 'max_temp', "total_rain", "gust_speed_km_h")]
colnames(selected_data_0) <- c("Wind Speed (km/h)", "Station Pressure (hPa)", "Snow (cm)", 
                               "Min Temperature (°C)", "Log Mean Temperature (°F)", 
                               "Max Temperature (°C)", "Total Rain (mm)", 
                               "Gust Speed (km/h)")

cor_matrix <- cor(selected_data_0)
corrplot(cor_matrix, method = "color", type = "upper", 
         tl.col = "black", tl.srt = 45, 
         col = colorRampPalette(c("blue", "white", "red"))(200))

selected_data_1 <- analysis_data[, c("wind_speed", "pressure_station", "snow", "log_mean_temp", "total_rain", "gust_speed_km_h")]
colnames(selected_data_1) <- c("Wind Speed (km/h)", "Station Pressure (hPa)", "Snow (cm)", 
                               "Log Mean Temperature (°F)", "Total Rain (mm)", 
                               "Gust Speed (km/h)")

cor_matrix <- cor(selected_data_1)
corrplot(cor_matrix, method = "color", type = "upper", 
         tl.col = "black", tl.srt = 45, 
         col = colorRampPalette(c("blue", "white", "red"))(200))

# Scatter plot with linear trendline for Max Temperature vs. Mean Temperature
ggplot(analysis_data, aes(x = `max_temp`, y = `mean_temp`)) +
  geom_point(color = "blue", alpha = 0.6) +   # Scatter plot
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add linear trendline
  labs(title = "Correlation: Max Temperature vs. Mean Temperature",
       x = "Max Temperature (°C)",
       y = "Mean Temperature (°C)") +
  theme_minimal()

# Scatter plot with linear trendline for Min Temperature vs. Mean Temperature
ggplot(analysis_data, aes(x = `min_temp`, y = `mean_temp`)) +
  geom_point(color = "green", alpha = 0.6) +   # Scatter plot
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add linear trendline
  labs(title = "Correlation: Min Temperature vs. Mean Temperature",
       x = "Min Temperature (°C)",
       y = "Mean Temperature (°C)") +
  theme_minimal()
```

@@

```{r}
fitted_values <- fitted(m1)
ggplot(data = train_data, aes(x = fitted_values, y = mean_temp_F)) +
  geom_point(color = "blue", alpha = 0.7) +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed", size = 1) +
  labs(x = "Predicted Temperature", y = "Actual Temperature", title = "Actual Temperature vs. Predicted Values")

@@

```{r}
fitted_values <- fitted(m2)
ggplot(data = train_data, aes(x = fitted_values, y = mean_temp_F)) +
  geom_point(color = "blue", alpha = 0.7) +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed", size = 1) +
  labs(x = "Predicted Temperature", y = "Actual Temperature", title = "Actual Temperature vs. Predicted Values")
```

@@

\appendix
# Appendix {-}

@@

# Summary Statistic of cleaned dataset
\newpage
```{r}
#| label: tbl-summary
#| tbl-cap: Summary Statistics of Raw Climate Data 
#| echo: false
#| warning: false
# Generate the table
analysis_data |>
  select(wind_speed, total_precipitation, snow, pressure_station,
  max_temp, min_temp) |>
  summary() |>
  kable(
    col.names = c(
      "Wind Speed", "Total Precipitation", "Snow", "Pres.", 
      "Max Temp", "Min Temp"
    ),
    booktabs = TRUE,
    align = "c",  # Center-align all columns
    format = "latex"  # Ensure LaTeX output
  )
analysis_data |>
  select(mean_temp, total_rain, gust_speed_km_h,
  mean_temp_F, log_mean_temp) |>
  summary() |>
  kable(
    col.names = c(
      "Mean Temp", "Rain", 
      "Max Gust Speed", "Mean Temp in F", "Log of Mean Temp"
    ),
    booktabs = TRUE,
    align = "c",  # Center-align all columns
    format = "latex"  # Ensure LaTeX output
  )
```
\newpage
